Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-11-27 GWLDCII.tex
specified citations in paper 2

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrands associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

@@

The first condition bounds the total number of propagators in the diagram, while the second limits how densely propagators can be fitted into any part of the diagram; in particular, the second condition prohibits propagators that connect two adjacent edges.  Note that since $\cP$ is a set, we may not have two propagators starting and ending on the same pair of edges. The third condition simply requires that propagators are non-crossing.

These conditions are imposed by the physical interpretation of Wilson loop diagrams (see for example \cite{Adamo:2011pv,Adamo:2012xe,wilsonloop,LipsteinMason}). We therefore restrict our attention in this paper to admissible Wilson loop diagrams and their subdiagrams.

@@

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, a diagram associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite{Arkani-Hamed:2013jha}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting. This integral is built from two main components: an interaction matrix $C(W)$, and a volume form $\Omega(W)$. These are described next.

@@

The matrix $C(W)$ defines a positroid cell associated to $W$ in the positive Grassmannian $\Gr(n, |\cP|)$, which we denote by $\Sigma(W)$ \cite{wilsonloop}. Marcott \cite{WLDdim} shows that  the matrix $C(W)$ parametrizes a space whose closure is exactly the closure of $\Sigma(W)$. That is, the space defined by $C(W)$ and the cell $\Sigma(W)$ agree up to a set of measure $0$. Most of the work in this paper and in the previous paper in the series \cite{generalcombinatoricsI} focuses on characterizing the positroid cell associated to each Wilson loop diagram $W$.

@@

One may also interpret the functional $I(W)$ as associating a volume to each cell $\Sigma(W)$ \cite{wilsonloop, Amplituhedronsquared, HeslopStewart}, where the volume is determined by the external particles involved. Specifically, the integrand is written in terms of the elements of $C(W)$ as  \begin{equation} \label{eq:omega(W) def}\Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V(p_r)} \textrm{d}x_{p_r}}{R(W)} \;, \end{equation} multiplied by a $\delta$ function that evaluates the matrix $C(W)$ on the external particles. The denominator $R(W)$ is given in Definition~\ref{def R(W)} below. For the purposes of this paper, we restrict our attention to $\Omega(W)$ only, and ignore the $\delta$ function in the integrand of $I(W)$. The interested reader is referred to \cite{Adamo:2012xe,HeslopStewart,LipsteinMason} for more information about the integrand of $I(W)$.

For a fixed number of propagators and particles, the tree level amplitude of the Wilson loop diagrams is computed as the sum of the $I(W)$ for all $W$ of the correct size. By \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha} this sum is finite for all sets of external particles, i.e.
\bas \mathcal{A}_{k, n}^{tree} = \sum_{\substack{W = (\cP, [n]), \\ |\cP| = k}} I(W) <\infty. \eas

@@

We summarize here only the portion of positroid theory that we require in this paper; the interested reader is referred to \cite{Postnikov} for more details.

@@

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces and Le diagrams, which we discuss. In order to do so, we first need some preliminary definitions.

For each $j \in [n]$, we can define a total order $<_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

@@

An algorithm for constructing the Grassmann necklace of a Le diagram also exists; this was given by Oh in \cite{Oh}. A method for using the Le diagram to test whether a given $k$-subset is a basis for the corresponding positroid or not was given by Casteels in \cite{CasteelsPaths}.

@@

However, to go from the Wilson loop diagram $W$ to the positroid cell, one had to go through the matroid explicitly realized via $C(W)$, then make a list of all its nonzero $k\times k$ minors and extract the Le diagram or Grassmann necklace via the methods described in Section~\ref{sec:positroid background}. In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible Wilson loop diagrams to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

\begin{rmk}
While this paper primarily approaches proofs from a combinatorial point of view, readers with a background in matroid theory are encouraged to keep the matroidal viewpoint in mind throughout; we expect that the two approaches will be complementary and will build upon each other. See the first paper in this series \cite{generalcombinatoricsI} for some related results concerning the positroids of Wilson loop diagrams obtained by viewing them primarily as matroids.
\end{rmk}

@@

Note that the positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite{non-orientable}, Agarwala and Marcott associate a Deodhar component\footnote{Since Deodhar components play no role in this paper, the interested reader is referred to \cite{non-orientable} for more details on this.} to each Wilson loop diagram, and show that the dimension of this Deodhar component is the sum of $|\cP|$ and the dimension of the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
Combining this with Theorem~\ref{thm dim}, we see that the Deodhar component associated to a Wilson loop diagram is $4|\cP|$-dimensional, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

Marcott in \cite{WLDdim} gives a different proof of the $3|\cP|$-dimensionality of $M(W)$ which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  It is worth noting that while the Amplituhedon is also defined by a projection of a set of $4|\cP|$ dimensional positroid cells, there is no evidence to suggest that the $3|\cP|$ dimensional cells of $\Gr(|\cP|, n)$ defiining the Wilson loop diagrams are related to the $4|\cP|$ dimensional cells defining the Amplituhedron.  \sanote{This para adresses point 3 of report Y. Should there be a similar paragraph in the intro? Should it be moved?}

@@

Since SYM $N=4$ is a finite theory \cite{Alday:2007hr, ParkeTaylorids}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite. That is, for a fixed set of external particles $\cZ$ we must have
\bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas
This holds if all poles of the integrands cancel. As a step towards this, in this section we prove some results about the structure of the denominator $R(W)$ of the integrand, as defined in Definition~\ref{def R(W)}.

>>>>

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite[section 11]{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrands associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

@@

The first condition bounds the total number of propagators in the diagram, and comes from the fact that these diagrams represent $N^k$MHV interctions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{adamo:2011pr} while the second limits how densely propagators can be fitted into any part of the diagram; in particular, the second condition prohibits propagators that connect two adjacent edges. This can be derived from the integrals associated to the diagrams, as in. Note that since $\cP$ is a set, we may not have two propagators starting and ending on the same pair of edges. \cite[Theorem 1.13]{wilsonloop}. The third condition simply requires that propagators are non-crossing. This follows from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipstenMason}

@@

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, a diagram associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite[section 2]{Amplituhedronsquared}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting. This integral is built from two main components: an interaction matrix $C(W)$, and a volume form $\Omega(W)$. These are described next.

@@

The matrix $C(W)$ defines a positroid cell associated to $W$ in the positive Grassmannian $\Gr(n, |\cP|)$, which we denote by $\Sigma(W)$ \cite[Theorem 3.41]{wilsonloop}. Marcott \cite[Theorem 8.4]{WLDdim} shows that  the matrix $C(W)$ parametrizes a space whose closure is exactly the closure of $\Sigma(W)$. That is, the space defined by $C(W)$ and the cell $\Sigma(W)$ agree up to a set of measure $0$. Most of the work in this paper and in the previous paper in the series \cite{generalcombinatoricsI} focuses on characterizing the positroid cell associated to each Wilson loop diagram $W$.

@@

One may also interpret the functional $I(W)$ as associating a volume to each cell $\Sigma(W)$ \cite{wilsonloop, Amplituhedronsquared, HeslopStewart}, where the volume is determined by the external particles involved. Specifically, the integrand is written in terms of the elements of $C(W)$ as  \begin{equation} \label{eq:omega(W) def}\Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V(p_r)} \textrm{d}x_{p_r}}{R(W)} \;, \end{equation} multiplied by a $\delta$ function that evaluates the matrix $C(W)$ on the external particles. The denominator $R(W)$ is given in Definition~\ref{def R(W)} below. For the purposes of this paper, we restrict our attention to $\Omega(W)$ only, and ignore the $\delta$ function in the integrand of $I(W)$. The interested reader is referred to \cite{Adamo:2012xe,HeslopStewart,LipsteinMason} for more information about the integrand of $I(W)$. \sanote{not putting sections here, but I'd like a second thought}

For a fixed number of propagators and particles, the tree level amplitude of the Wilson loop diagrams is computed as the sum of the $I(W)$ for all $W$ of the correct size. \hlfix{By {\cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}} this sum is finite for all sets of external particles,}{fix this citation} i.e.
\bas \mathcal{A}_{k, n}^{tree} = \sum_{\substack{W = (\cP, [n]), \\ |\cP| = k}} I(W) <\infty. \eas

@@

We summarize here only the portion of positroid theory that we require in this paper; the interested reader is referred to \cite[sections 6, 16]{Postnikov} for more details.

@@

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces \cite[section 16]{Postnikov} and Le diagrams\cite[section 6]{Postnikov}, which we discuss. In order to do so, we first need some preliminary definitions.

For each $j \in [n]$, we can define a total order $<_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}) \sanote{Not citing a thm here either as we give the dfn}, which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

@@

An algorithm for constructing the Grassmann necklace of a Le diagram also exists; this was given by Oh in \cite[section 4]{Oh}. A method for using the Le diagram to test whether a given $k$-subset is a basis for the corresponding positroid or not was given by Casteels in \cite{CasteelsPaths}.

@@

However, to go from the Wilson loop diagram $W$ to the positroid cell, one had to go through the matroid explicitly realized via $C(W)$, then make a list of all its nonzero $k\times k$ minors and extract the Le diagram or Grassmann necklace via the methods described in Section~\ref{sec:positroid background}. In \cite[section 4]{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible Wilson loop diagrams to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

\begin{rmk}
While this paper primarily approaches proofs from a combinatorial point of view, readers with a background in matroid theory are encouraged to keep the matroidal viewpoint in mind throughout; we expect that the two approaches will be complementary and will build upon each other. See the first paper in this series \cite[section 3]{generalcombinatoricsI} for some related results concerning the positroids of Wilson loop diagrams obtained by viewing them primarily as matroids.
\end{rmk}

@@

Note that the positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite[section 2]{non-orientable}, Agarwala and Marcott associate a Deodhar component\footnote{Since Deodhar components play no role in this paper, the interested reader is referred to \cite{non-orientable} for more details on this.} to each Wilson loop diagram, and show that the dimension of this Deodhar component is the sum of $|\cP|$ and the dimension of the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
Combining this with Theorem~\ref{thm dim}, we see that the Deodhar component associated to a Wilson loop diagram is $4|\cP|$-dimensional, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

Marcott in \cite[Theorem 8.4]{WLDdim} gives a different proof of the $3|\cP|$-dimensionality of $M(W)$ which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  It is worth noting that while the Amplituhedon is also defined by a projection of a set of $4|\cP|$ dimensional positroid cells, there is no evidence to suggest that the $3|\cP|$ dimensional cells of $\Gr(|\cP|, n)$ defiining the Wilson loop diagrams are related to the $4|\cP|$ dimensional cells defining the Amplituhedron.  \sanote{This para adresses point 3 of report Y. Should there be a similar paragraph in the intro? Should it be moved?}

@@

\hlfix{Since SYM $N=4$ is a finite theory {\cite{Alday:2007hr, ParkeTaylorids}}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite.}{Fix citations} That is, for a fixed set of external particles $\cZ$ we must have
\bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas
This holds if all poles of the integrands cancel. As a step towards this, in this section we prove some results about the structure of the denominator $R(W)$ of the integrand, as defined in Definition~\ref{def R(W)}.
